---
title: "Data 621 Homework 5"
author: "Trishita Nath"
date: "12/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

In this homework assignment, you will explore, analyze and model a data set containing information on approximately 12,000 commercially available wines. The variables are mostly related to the chemical properties of the wine being sold. The response variable is the number of sample cases of wine that were purchased by wine distribution companies after sampling a wine. These cases would be used to provide tasting samples to restaurants and wine stores around the United States. The more sample cases purchased, the more likely is a wine to be sold at a high end restaurant. A large wine manufacturer is studying the data in order to predict the number of wine cases ordered based upon the wine characteristics. If the wine manufacturer can predict the number of cases, then that manufacturer will be able to adjust their wine offering to maximize sales.

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=FALSE, results='hide'}
library(dplyr)
library(tidyverse)
library(kableExtra)
library(visdat)
library(DT)
library(psych)
library(corrplot)
library(qqplotr)
library(MASS)
library(matrixStats)
library(pacman)
library(bestglm)
library(glmnet)
library(AICcmodavg)
library(RcmdrMisc)
library(caret)
library(mice)
p_load(Hmisc, xlsx, xtable, knitr, scales, magrittr, tidyverse, stringr, e1071, corrplot, knitcitations, bibtex, missForest, abc,
       foreach, doParallel, stargazer, forecast, matrixStats, glmulti, leaps, data.table, highlight, car, Amelia, caret)
```

# DATA EXPLORATION

## Loading Data

```{r Import Dataset, message=FALSE, warning=FALSE, echo=FALSE, results='hide'}
training_dataset <- read_csv("https://raw.githubusercontent.com/nathtrish334/Data-621/main/HW5/wine-training-data.csv") %>%
  dplyr::select(TARGET, everything())
evaluation_dataset <- read_csv('https://raw.githubusercontent.com/nathtrish334/Data-621/main/HW5/wine-evaluation-data.csv')
```

### Training Dataset
```{r Training Dataset, message=FALSE, warning=FALSE, echo=FALSE, results='show'}
# Show Entire dataset
DT::datatable(training_dataset)
```
The training dataset has 14 predictors, 1 response variable and 12795 observations.

### Evaluation Dataset
```{r Evaluation Dataset, message=FALSE, warning=FALSE, echo=FALSE, results='show'}
# Show Entire dataset
DT::datatable(evaluation_dataset)
```

## Missing Values & Data Type Check

From the 14 predictor variables, 3 are categorical (`LabelAppeal`,`AcidIndex`,`STARS`), the other 11 are continuous numerical. The response variable `TARGET` is categorical.

```{r, glimpse training set, fig.cap = "Glimpse of `training set`", message=FALSE, warning=FALSE, echo=FALSE, results='show'}
glimpse(training_dataset)
```

```{r, glimpse evaluation set, fig.cap= "Glimpse of `evaluation set`", message=FALSE, warning=FALSE, echo=FALSE, results='show'}
glimpse(evaluation_dataset)
```

```{r, missing values & data type check, fig.cap= "Missing Values & Data Type Check", message=FALSE, warning=FALSE, echo=FALSE, results='show'}
library(gridExtra)
p_t_dt <- vis_dat(training_dataset)
p_t_m <- vis_miss(training_dataset)
p_e_dt <- vis_dat(evaluation_dataset)
p_e_m <- vis_miss(evaluation_dataset)
grid.arrange(p_t_m,p_e_m, p_t_dt,p_e_dt,ncol = 2, 
             widths = c(1,1),
             heights = c(1.5,1),
             top = 'Missing Values & Data Type Check
             Training Set                                                     Evaluation Set')
#Amelia::missmap(training_dataset, main = "Missing vs Observed Values in Traning Data")
Non_NAs <- sapply(training_dataset, function(y) sum(length(which(!is.na(y)))))
NAs <- sapply(training_dataset, function(y) sum(length(which(is.na(y)))))
NA_Percent <- NAs / (NAs + Non_NAs)
NA_SUMMARY <- data.frame(Non_NAs,NAs,NA_Percent)
Amelia::missmap(training_dataset, main = "Missing vs Observed Values in Traning Data")
kable(NA_SUMMARY)
```

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='show'}
temp <- training_dataset %>%
  #select(TARGET, STARS, Alcohol, ResidualSugar, Chlorides, FreeSulfurDioxide, TotalSulfurDioxide, pH, Sulphates) %>%
  gather("variable", "value", -TARGET) %>%
  mutate(na = ifelse(is.na(value), "Yes", "No")) %>%
  group_by(TARGET, na, variable) %>%
  tally() %>%
  ungroup()
 temp <- temp %>%
    group_by(variable, na) %>%
    summarise(total = sum(n)) %>%
    merge(temp) %>%
    mutate(share = n / total)
 
 ggplot(temp, aes(TARGET, share, fill = na)) +
   geom_bar(stat = "identity", position = "dodge") +
   scale_fill_brewer(palette = "Set1") +
   facet_wrap(~variable, ncol = 4) + 
   ylab("Percent of Group")
```

### Summary statistics for the training dataset

A binary logistic regression model is built using the `training set`, hence the `training set` is used for data exploration.

```{r summary, fig.cap= "`training set`", message=FALSE, warning=FALSE, echo=FALSE, results='show'}
training_modified <- training_dataset %>% 
  dplyr::select(-(INDEX)) %>% 
  dplyr::select(TARGET, everything())
DT::datatable(training_modified)
```


```{r statistics summary}
summary(training_modified)
```

## Outliers

```{r Box Plot on training set, fig.cap = "Boxplot: Scaled Training Set", warning=FALSE, message=FALSE}
training_modified %>%
  scale() %>%
  as.data.frame() %>%
  stack() %>%
  ggplot(aes(x = ind, y = values)) +
  geom_boxplot(fill = 'deeppink4') +
  labs(title = 'Boxplot: Scaled Training Set',
       x = 'Variables',
       y = 'Normalized_Values')+
  theme(panel.background = element_rect(fill = 'grey'),axis.text.x=element_text(size=10, angle=90))  
```

From the above Boxplot outliers occur in: `FixedAcidity`, `VolatileAcidity`, `CitricAcid`, `ResidualSugar`, `Chlorides`, `FreeSulfurDioxide`, `TotalSulfurDioxide`, `Density`, `pH`, `Sulphates`, `Alcohol`, `LabelAppeal` and`AcidIndex`.

## Histograms

- **FixedAcidity:**

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
library(qqplotr)
with(training_modified, c(summary(FixedAcidity), SD=sd(FixedAcidity), Skew=skewness(FixedAcidity), Kurt=kurtosis(FixedAcidity)))
hist <- ggplot(training_modified, aes(FixedAcidity)) + geom_histogram(fill = 'dodgerblue', binwidth = 4, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of FixedAcidity') + theme(plot.title = element_text(hjust = 0.5)) 
qq_plot <- ggplot(training_modified, aes(sample=FixedAcidity)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of FixedAcidity") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 
box_plot <- ggplot(training_modified, aes(x="", FixedAcidity)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of FixedAcidity', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()
box_TARGET <- ggplot(training_modified, aes(x=factor(TARGET), FixedAcidity)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='TARGET', title = 'Boxplot of FixedAcidity by TARGET') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
grid.arrange(hist, qq_plot, box_plot, box_TARGET, ncol=2)
```

- **VolatileAcidity:**

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
with(training_modified, c(summary(VolatileAcidity), SD=sd(VolatileAcidity), Skew=skewness(VolatileAcidity), Kurt=kurtosis(VolatileAcidity)))
hist <- ggplot(training_modified, aes(VolatileAcidity)) + geom_histogram(fill = 'dodgerblue', binwidth = .5, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of VolatileAcidity') + theme(plot.title = element_text(hjust = 0.5)) 
qq_plot <- ggplot(training_modified, aes(sample=VolatileAcidity)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of VolatileAcidity") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 
box_plot <- ggplot(training_modified, aes(x="", VolatileAcidity)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of VolatileAcidity', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()
box_TARGET <- ggplot(training_modified, aes(x=factor(TARGET), VolatileAcidity)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='TARGET', title = 'Boxplot of VolatileAcidity by TARGET') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
grid.arrange(hist, qq_plot, box_plot, box_TARGET, ncol=2)
```

- **CitricAcid:**

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
with(training_modified, c(summary(CitricAcid), SD=sd(CitricAcid), Skew=skewness(CitricAcid), Kurt=kurtosis(CitricAcid)))
hist <- ggplot(training_modified, aes(CitricAcid)) + geom_histogram(fill = 'dodgerblue', binwidth = 1, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of CitricAcid') + theme(plot.title = element_text(hjust = 0.5)) 
qq_plot <- ggplot(training_modified, aes(sample=CitricAcid)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of CitricAcid") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 
box_plot <- ggplot(training_modified, aes(x="", CitricAcid)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of CitricAcid', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()
box_TARGET <- ggplot(training_modified, aes(x=factor(TARGET), CitricAcid)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='TARGET', title = 'Boxplot of CitricAcid by TARGET') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
grid.arrange(hist, qq_plot, box_plot, box_TARGET, ncol=2)
```

- **ResidualSugar:**

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
with(training_modified, c(summary(ResidualSugar), SD=sd(ResidualSugar), Skew=skewness(ResidualSugar), Kurt=kurtosis(ResidualSugar)))
hist <- ggplot(training_modified, aes(ResidualSugar)) + geom_histogram(fill = 'dodgerblue', binwidth = 20, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of ResidualSugar') + theme(plot.title = element_text(hjust = 0.5)) 
qq_plot <- ggplot(training_modified, aes(sample=ResidualSugar)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of ResidualSugar") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 
box_plot <- ggplot(training_modified, aes(x="", ResidualSugar)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of ResidualSugar', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()
box_TARGET <- ggplot(training_modified, aes(x=factor(TARGET), ResidualSugar)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='TARGET', title = 'Boxplot of ResidualSugar by TARGET') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
grid.arrange(hist, qq_plot, box_plot, box_TARGET, ncol=2)
```

- **Chlorides:**

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
with(training_modified, c(summary(Chlorides), SD=sd(Chlorides), Skew=skewness(Chlorides), Kurt=kurtosis(Chlorides)))
hist <- ggplot(training_modified, aes(Chlorides)) + geom_histogram(fill = 'dodgerblue', binwidth = .2, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of Chlorides') + theme(plot.title = element_text(hjust = 0.5)) 
qq_plot <- ggplot(training_modified, aes(sample=Chlorides)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of Chlorides") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 
box_plot <- ggplot(training_modified, aes(x="", Chlorides)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of Chlorides', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()
box_TARGET <- ggplot(training_modified, aes(x=factor(TARGET), Chlorides)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='TARGET', title = 'Boxplot of Chlorides by TARGET') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
grid.arrange(hist, qq_plot, box_plot, box_TARGET, ncol=2)
```

- **FreeSulfurDioxide**

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
with(training_modified, c(summary(FreeSulfurDioxide), SD=sd(FreeSulfurDioxide), Skew=skewness(FreeSulfurDioxide), Kurt=kurtosis(FreeSulfurDioxide)))
hist <- ggplot(training_modified, aes(FreeSulfurDioxide)) + geom_histogram(fill = 'dodgerblue', binwidth = 50, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of FreeSulfurDioxide') + theme(plot.title = element_text(hjust = 0.5)) 
qq_plot <- ggplot(training_modified, aes(sample=FreeSulfurDioxide)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of FreeSulfurDioxide") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 
box_plot <- ggplot(training_modified, aes(x="", FreeSulfurDioxide)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of FreeSulfurDioxide', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()
box_TARGET <- ggplot(training_modified, aes(x=factor(TARGET), FreeSulfurDioxide)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='TARGET', title = 'Boxplot of FreeSulfurDioxide by TARGET') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
grid.arrange(hist, qq_plot, box_plot, box_TARGET, ncol=2)
```

- **TotalSulfurDioxide**

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
with(training_modified, c(summary(TotalSulfurDioxide), SD=sd(TotalSulfurDioxide), Skew=skewness(TotalSulfurDioxide), Kurt=kurtosis(TotalSulfurDioxide)))
hist <- ggplot(training_modified, aes(TotalSulfurDioxide)) + geom_histogram(fill = 'dodgerblue', binwidth = 200, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of TotalSulfurDioxide') + theme(plot.title = element_text(hjust = 0.5)) 
qq_plot <- ggplot(training_modified, aes(sample=TotalSulfurDioxide)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of TotalSulfurDioxide") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 
box_plot <- ggplot(training_modified, aes(x="", TotalSulfurDioxide)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of TotalSulfurDioxide', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()
box_TARGET <- ggplot(training_modified, aes(x=factor(TARGET), TotalSulfurDioxide)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='TARGET', title = 'Boxplot of TotalSulfurDioxide by TARGET') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
grid.arrange(hist, qq_plot, box_plot, box_TARGET, ncol=2)
```

- **Density:**

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
with(training_modified, c(summary(Density), SD=sd(Density), Skew=skewness(Density), Kurt=kurtosis(Density)))
hist <- ggplot(training_modified, aes(Density)) + geom_histogram(fill = 'dodgerblue', binwidth = .05, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of Density') + theme(plot.title = element_text(hjust = 0.5)) 
qq_plot <- ggplot(training_modified, aes(sample=Density)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of Density") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 
box_plot <- ggplot(training_modified, aes(x="", Density)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of Density', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()
box_TARGET <- ggplot(training_modified, aes(x=factor(TARGET), Density)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='TARGET', title = 'Boxplot of Density by TARGET') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
grid.arrange(hist, qq_plot, box_plot, box_TARGET, ncol=2)
```

- **Sulphates:**

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
with(training_modified, c(summary(Sulphates), SD=sd(Sulphates), Skew=skewness(Sulphates), Kurt=kurtosis(Sulphates)))
hist <- ggplot(training_modified, aes(Sulphates)) + geom_histogram(fill = 'dodgerblue', binwidth = .5, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of Sulphates') + theme(plot.title = element_text(hjust = 0.5)) 
qq_plot <- ggplot(training_modified, aes(sample=Sulphates)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of Sulphates") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 
box_plot <- ggplot(training_modified, aes(x="", Sulphates)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of Sulphates', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()
box_TARGET <- ggplot(training_modified, aes(x=factor(TARGET), Sulphates)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='TARGET', title = 'Boxplot of Sulphates by TARGET') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
grid.arrange(hist, qq_plot, box_plot, box_TARGET, ncol=2)
```

- **Alcohol:**

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
with(training_modified, c(summary(Alcohol), SD=sd(Alcohol), Skew=skewness(Alcohol), Kurt=kurtosis(Alcohol)))
hist <- ggplot(training_modified, aes(Alcohol)) + geom_histogram(fill = 'dodgerblue', binwidth = 2, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of Alcohol') + theme(plot.title = element_text(hjust = 0.5)) 
qq_plot <- ggplot(training_modified, aes(sample=Alcohol)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of Alcohol") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 
box_plot <- ggplot(training_modified, aes(x="", Alcohol)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of Alcohol', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()
box_TARGET <- ggplot(training_modified, aes(x=factor(TARGET), Alcohol)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='TARGET', title = 'Boxplot of Alcohol by TARGET') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
grid.arrange(hist, qq_plot, box_plot, box_TARGET, ncol=2)
```

- **LabelAppeal:**

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
with(training_modified, c(summary(LabelAppeal), SD=sd(LabelAppeal), Skew=skewness(LabelAppeal), Kurt=kurtosis(LabelAppeal)))
hist <- ggplot(training_modified, aes(LabelAppeal)) + geom_histogram(fill = 'dodgerblue', binwidth = 1, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of LabelAppeal') + theme(plot.title = element_text(hjust = 0.5)) 
qq_plot <- ggplot(training_modified, aes(sample=LabelAppeal)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of LabelAppeal") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 
box_plot <- ggplot(training_modified, aes(x="", LabelAppeal)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of LabelAppeal', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()
box_TARGET <- ggplot(training_modified, aes(x=factor(TARGET), LabelAppeal)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='TARGET', title = 'Boxplot of LabelAppeal by TARGET') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
grid.arrange(hist, qq_plot, box_plot, box_TARGET, ncol=2)
```

- **AcidIndex:**

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
with(training_modified, c(summary(AcidIndex), SD=sd(AcidIndex), Skew=skewness(AcidIndex), Kurt=kurtosis(AcidIndex)))
hist <- ggplot(training_modified, aes(AcidIndex)) + geom_histogram(fill = 'dodgerblue', binwidth = 2, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of AcidIndex') + theme(plot.title = element_text(hjust = 0.5)) 
qq_plot <- ggplot(training_modified, aes(sample=AcidIndex)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of AcidIndex") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 
box_plot <- ggplot(training_modified, aes(x="", AcidIndex)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of AcidIndex', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()
box_TARGET <- ggplot(training_modified, aes(x=factor(TARGET), AcidIndex)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='TARGET', title = 'Boxplot of AcidIndex by TARGET') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
grid.arrange(hist, qq_plot, box_plot, box_TARGET, ncol=2)
```

- **STARS:**

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
with(training_modified, c(summary(STARS), SD=sd(STARS), Skew=skewness(STARS), Kurt=kurtosis(STARS)))
hist <- ggplot(training_modified, aes(STARS)) + geom_histogram(fill = 'dodgerblue', binwidth = 1, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of STARS') + theme(plot.title = element_text(hjust = 0.5)) 
qq_plot <- ggplot(training_modified, aes(sample=STARS)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of STARS") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 
box_plot <- ggplot(training_modified, aes(x="", STARS)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of STARS', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()
box_TARGET <- ggplot(training_modified, aes(x=factor(TARGET), STARS)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='TARGET', title = 'Boxplot of STARS by TARGET') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
grid.arrange(hist, qq_plot, box_plot, box_TARGET, ncol=2)
```

## Density Plot

```{r, fig.cap = "Density Plot: Training Set", eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
training_modified %>%
  select_if(is.numeric) %>%
  keep(is.numeric) %>%                     # Keep only numeric columns
  gather() %>%                             # Convert to key-value pairs
  ggplot(aes(x=value)) +                   # Plot the values
    facet_wrap(~key, scales = "free") +    # In separate panels
    geom_density()  
```

From the scaled histogram and density plots, `AcidIndex` is skewed right; `AcidIndex`, `STARS`, `LabelAppeal` and `TARGET` have multimodal distribution; while other variables seem to be normally distrbuted. 

## Correlation Plot

Correlation between variables in the dataset. 

```{r correlation hitmap, fig.cap = "Correlation Pie Chart: Training Set"}
drop_na(training_modified) %>%
  select_if(is.numeric) %>%
  cor() %>%
  #corrplot(method = "square", type = "upper", order = 'hclust', tl.col = "black", diag = FALSE, bg= 'white', col = colorRampPalette(c('deeppink4','white','steelblue1'))(100))
  corrplot.mixed(upper = 'pie', lower = 'number', order = 'hclust', tl.col = "black") 
```

From the correlation matrix below, the response variable `TARGET` has strong positive relationship with variables `FixedAcidity`,`CitricAcid`,`ResidualSugar`,`Density`,`Alcohol`. 

```{r include=TRUE, fig.cap = "Correlation Chart: Training Set", echo=TRUE, warning=FALSE, message=FALSE}
PerformanceAnalytics::chart.Correlation(training_dataset, histogram=TRUE, pch=19)
```

## Scatter plots

```{r}
training_dataset %>%
  gather(-TARGET, key = "key", value = "ResponseVariables") %>%
  ggplot(aes(x = ResponseVariables, y = TARGET)) +
  geom_point(size = .5) +
  geom_smooth(method='lm',formula=y~x, color = 'dark grey')+
  facet_wrap(~ key, scales = "free")+
  ggthemes::theme_tufte()+
  ylab('Cases Bought')
```

There are no unusual patterns. `STARS` and `LableAppleal` seem to have the greatest correlation.


```{r test}
training_dataset %>%
  dplyr::select(-(INDEX)) %>%
  cor() %>%
  as.data.frame() %>%
  rownames_to_column('Variable') %>%
  dplyr::rename(Correlation_vs_Response = TARGET)
```

## Data Exploration Summary

The data exploration process can be summarized in the data dictionary below:

```{r data dictionary, fig.cap = "Data Dictionary"}
data_stat <- training_dataset %>% 
  dplyr::select(-TARGET,-INDEX) %>%
  gather() %>%
  group_by(key) %>%
  summarise(Mean = mean(value),
            Median = median(value),
            Max = max(value),
            Min = min(value),
            SD = sd(value))
data_cor <- training_dataset %>%
  cor() %>%
  as.data.frame() %>% 
  dplyr::select(TARGET) %>% 
  rownames_to_column('Variable') %>%
  dplyr::rename(Correlation_vs_Response = TARGET)
training_dataset %>% 
  gather() %>%
  dplyr::select(key) %>%
  unique() %>%
  dplyr::rename(Variable = key) %>%
  mutate(
         Missing_Value = 'No') %>%
  left_join(data_stat, by = c('Variable'='key')) %>%
  left_join(data_cor, by = 'Variable') %>%
  mutate_if(is.numeric,round,2) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),full_width = F)
```

# DATA PREPARATION

I split the data into training and test.

The MICE package comes in handy.

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
set.seed(999) 
sampl = caTools::sample.split(training_modified$TARGET, SplitRatio = .80)
train_dataset_1 <- subset(training_modified, sampl == TRUE)
test_dataset_1 <- subset(training_modified, sampl == FALSE)
train_dataset_2 <-  as.data.frame(tidyr::complete(mice(train_dataset_1, m=1, maxit = 5, seed = 42)))
test_dataset_2 <- as.data.frame(tidyr::complete(mice(test_dataset_1, m=1, maxit = 5, seed = 42)))
```

Perform log transform to the data. 

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
train_dataset_2$AcidIndex <- log(train_dataset_2$AcidIndex)
test_dataset_2$AcidIndex <- log(test_dataset_2$AcidIndex)
```

# BUILD MODELS

## Poisson Models

### Model 1: Poisson Model without imputations

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
require(ggplot2)
require(gridExtra)
model1 = glm(TARGET ~  ., data=train_dataset_1, family=poisson)
summary(model1)
plot(model1)
grid.arrange(hist, qq_plot, box_plot, box_TARGET, ncol=2)
```

### Model 2: Poisson Model  with imputations

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
model2 = glm(TARGET ~  ., data=train_dataset_2, family=poisson)
summary(model2)
plot(model2)
```

## Negative Binomial Models

### Model 3 : Negative Binomial without imputations

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
model3 <- glm.nb(TARGET ~ ., data = train_dataset_1)
summary(model3)
plot(model3)
```

### Model 4 : Negative Binomial with imputations

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
model4 <- glm.nb(TARGET ~ ., data = train_dataset_2)
summary(model4)
plot(model4)
```

## Linear Models

### Model 5 : Linear Model with imputations

Using Linear Regression Model on imputed training data.

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
model5 <- lm(TARGET ~ ., data = train_dataset_2)
summary(model5)
plot(model5)
```

## Model 6 : Ordinal Logistic Regression

This regression uses ordered factors. I would expect this to be one of the top performers.

```{r}
polrDF <- train_dataset_2
polrDF$TARGET <- as.factor(polrDF$TARGET)
model6 <- polr(TARGET ~ ., data = polrDF, Hess=TRUE)
summary(model6)
```

## Model 7 : Zero inflation 

Zero inflation understands that some Poisson distrobutions are dominated by many zeros. As such it corrects for this. This is one of the most promissing ones because as we saw in our data exploration, there were more zeros, and then normally distributed data after that. 

```{r}
library(pscl)
model7 <- zeroinfl(TARGET ~ . | STARS, data = train_dataset_2, dist = 'negbin')
summary(model7)
scatterPreds <- predict(model7, train_dataset_2)
qplot(train_dataset_2$TARGET, scatterPreds, main = 'Predicted vs Actual') + ggthemes::theme_tufte()
residPlot <- scatterPreds - train_dataset_2$TARGET
qplot(train_dataset_2$TARGET, residPlot, main = 'Residuals') + ggthemes::theme_tufte()
```

# MODEL SELECTION

## Compare Models by MSE/AIC

```{r, echo=TRUE, warning=FALSE, message=FALSE}
aic1 <- model1$aic
aic2 <- model2$aic
aic3 <- model3$aic
aic4 <- model4$aic
aic5 <- model5$aic
aic6 <- model6$aic
aic7 <- model7$aic
mse1 <- mean((train_dataset_2$TARGET - predict(model1))^2)
mse2 <- mean((train_dataset_2$TARGET - predict(model2))^2)
mse3 <- mean((train_dataset_2$TARGET - predict(model3))^2)
mse4 <- mean((train_dataset_2$TARGET - predict(model4))^2)
mse5 <- mean((train_dataset_2$TARGET - predict(model5))^2)
mse6 <- mean((train_dataset_2$TARGET - predict(model6))^2)
mse7 <- mean((train_dataset_2$TARGET - predict(model7))^2)
compare_aic_mse <- matrix(c(mse1, mse2, mse3, mse4, mse5, mse6, mse7,
                            aic1, aic2, aic3, aic4, aic5, aic6,
                            aic7),nrow=7,ncol=2,byrow=TRUE)
rownames(compare_aic_mse) <- c("Model1","Model2","Model3","Model4","Model5","Model6","Model7")
colnames(compare_aic_mse) <- c("MSE","AIC")
compare_models <- as.data.frame(compare_models)
kable(compare_aic_mse)  %>% 
  kable_styling(full_width = T)
```

## Compare Models by Loss

Now lets see the output of the Models using test data

We will use the squared loss to validate the model. We will use the squared difference to select a model (MSE) from predictions on the training sets. (Lower numbers are better.)

```{r}
modelValidation <- function(mod){
  preds = predict(mod, test_dataset_2)
  diffMat = as.numeric(preds) - as.numeric(test_dataset_2$TARGET)
  diffMat = diffMat^2
  loss <- mean(diffMat)
  return(loss)
}
compare_models <- matrix(c(modelValidation(model1),modelValidation(model2),
                     modelValidation(model3),modelValidation(model4),
                     modelValidation(model5),modelValidation(model6),
                     modelValidation(model7)),
                         nrow=7,ncol=1,byrow=TRUE)
rownames(compare_models) <- c("Model1","Model2","Model3","Model4","Model5","Model6","Model7")
colnames(compare_models) <- c("Loss:")
compare_models <- as.data.frame(compare_models)
compare_models
```

The squared loss metric tells how accurate our model is without caring about confidence intervals among others. Based on this parameter, the Zero Poisson Inflation model is the most accurate. 

model7 (Zero Poisson Inflations) is the best choice overally due to the following reasons:
- least loss
- good MSE score
- good AIC score

## Prediction on Evaluation Data

In the same manner as before, lets impute and use log transformation for AcidIndex.

```{r}
evaluation_data_modified <- evaluation_dataset %>% dplyr::select(-(IN)) %>% mice(m=1, maxit = 5, seed = 42)
imputed_eval_data <- as.data.frame(complete(evaluation_data_modified))
imputed_eval_data$AcidIndex <- log(imputed_eval_data$AcidIndex)
imputed_eval_data$TARGET <- predict(model7, newdata=imputed_eval_data)
write.csv(imputed_eval_data,"predicted_data.csv", row.names=FALSE)
data_predicted_eval <- read_csv("predicted_data.csv")
DT::datatable(data_predicted_eval)
```

- **TARGET:** Predicted Number of Cases Purchased

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE, results='show'}
options(width=100)
round(with(data_predicted_eval, c(summary(TARGET), StdD=sd(TARGET), Skew=skewness(TARGET), Kurt=kurtosis(TARGET))),2)
```


```
